A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-11-30 bioc-workflow.Rnw
Added section on foi

It is also possible for users to use their own marker list with the
\texttt{addMarkers} function. The user needs to create a named vector
of marker localisation, or a create a csv file with two columns (one
for the protein names, one for the corresponding sub-cellular marker
annotation) and pass the vector or filename respectively to the
function. As previously mentioned, the protein names of these markers
must match some (but not necessarily all) of the \texttt{MSnSet}'s
feature names. See \texttt{?addMarkers} for more details.

We now visualise these annotations along the PCA plot using the
\texttt{plot2D} function and then use the \texttt{addLegend} function
to map the marker classes to the pre-defined colours. We also display
the data along the first and seventh PCs using the \texttt{dims}
argument. Note that in these calls to the \texttt{plot2D} function, we
have omitted the \texttt{fcol} argument and use of the default
\texttt{"markers"} feature variable to annotate the plot. We choose
to display PCs 1 and 7 to illustrate that while upper principal
components explain much less variability in the data (2.23\% for PC7,
as opposed to 48.41\% for PC1), we see that the mitochondrial (purple)
and peroxisome (dark blue) clusters can be differenciated, despite the
apparent overlap in the visualisation of the two first PCs
(Figure \ref{fig:plotmarkers}).

@@

>>>>

\subsection{*Adding user-defined marker lists}
It is also possible for users to use their own marker list with the
\texttt{addMarkers} function. The user needs to create a named vector
of marker localisation, or a create a csv file with two columns (one
for the protein names, one for the corresponding sub-cellular marker
annotation) and pass the vector or filename respectively to the
function. As previously mentioned, the protein names of these markers
must match some (but not necessarily all) of the \texttt{MSnSet}'s
feature names. See \texttt{?addMarkers} for more details.

\subsection{*Visualising markers}
Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
\texttt{plot2D} function and then use the \texttt{addLegend} function
to map the marker classes to the pre-defined colours. We also display
the data along the first and seventh PCs using the \texttt{dims}
argument. Note that in these calls to the \texttt{plot2D} function, we
have omitted the \texttt{fcol} argument and use of the default
\texttt{"markers"} feature variable to annotate the plot. We choose
to display PCs 1 and 7 to illustrate that while upper principal
components explain much less variability in the data (2.23\% for PC7,
as opposed to 48.41\% for PC1), we see that the mitochondrial (purple)
and peroxisome (dark blue) clusters can be differenciated, despite the
apparent overlap in the visualisation of the two first PCs
(Figure \ref{fig:plotmarkers}).

@@

\subsection*{Features of interest}
In addition to adding annotation using the \texttt{addMarkers} function one can store specific sets of proteins by using the \textit{Features of interest} infrastructure from the \Biocpkg{MSnbase} package. If users have a specific subset of proteins they wish to highlight in their data they would first create a \texttt{FeaturesOfInterest} object. For example, if we wanted to highlight a proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some of the proteins known to form part of the 13S condensin complex, we would create a first create a \texttt{FeaturesOfInterest} object then use the function \texttt{highlightOnPlot} to highlight their location on a PCA plot.

\begin{figure}[!ht]
  \centering
<<foi, fig.width=10, out.width=".8\\textwidth">>= 
prots <- c("Q8CG48", "Q8CG47", "Q8K2Z4", "Q8C156")
foi13s <- FeaturesOfInterest(description = "13S consensin proteins",
                             fnames = prots, 
                             object = hl)
foi13s

plot2D(hl)
addLegend(hl, cex = .6)
highlightOnPlot(hl, foi13s)
@  
  \caption{Highlighting protein features of interest}
  \label{fig:foi}
\end{figure}

Users can also create several sets of \texttt{FeaturesOfInterest} object and store them in a \texttt{FoICollection}.

It is also worthly of note that it is possible to search for a specific protein of interest by \texttt{featureNames} or using any identifying information found in the \texttt{fData} columns by using the search box on the \texttt{pRolocVis} application part of the \Biocpkg{pRolocGUI} package (see section on interactive visualisation). This can be handy for quickly searching and highlighting proteins on the fly, the disavanatge here is that proteins can only be searched for a one-by-one basis.
